---
title: "AFModulus_Flex"
author: "Selen Manioglu, Astrid Stubbusch, Audrey Yeo"
date: "1/11/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## include Python cede (https://rstudio.github.io/reticulate/)
knitr::knit_engines$set(python = reticulate::eng_python)
#use_python("/usr/local/bin/python3.9")
#use_python("/usr/bin/python")
# py$x would access an x variable created within Python from R
# r.x would access to x variable created within R from Python

# clear workspace
rm(list = ls())

# set working directory
setwd("~/Documents/Zurich_PhD/courses/2021_01_11_Compulsory_LSZGS SysBio_Computational Biology/git_folder/")

# load packages
library(tidyverse)
```

# Atomic Force Microscopy - From Topology to Stiffness (Modulus)

The git page of this project can be found here <https://github.com/audreyyeoCH/AFModulus_Flex>.


\newpage
## Import of AFM curves (force vs. separation distance) for each pixel

```{r import, echo = FALSE}
Curve_name <- "AFModulus_Flex/F_vs_t_curves/20200619_.005.pfc-4069_ForceCurveIndex_45647.spm - NanoScope Analysis_F_vs_t.txt"
F_vs_t_curve1 <- read.table(Curve_name, 
                            quote="\"", 
                            header = T,
                            comment.char="", 
                            stringsAsFactors=FALSE)
F_vs_t_curve1 <- as_tibble(F_vs_t_curve1)

```

#### Plot 1 curve
```{r plot, echo=FALSE}
ggplot(data = F_vs_t_curve1, mapping = aes(x = ms, y = pN)) +
  geom_point() +
  geom_line() +
  ggtitle(Curve_name) +
    labs(
      x = "Time [ms]",
      y = "Force [pN]"
    )

```
We will use the force signal between time 0.0 - 0.5 ms as well as 1.5 - 2.0 ms as **baseline**.

We will use a sliding window approach to approximate the **gradient** of the linear slope within the time 0.5 - 1.2 ms.


\newpage
## Extract **maximal force (F max)** from each graph

```{r Fmax, echo=F}
Fmax       <- max(F_vs_t_curve1$pN)
Fmax_t     <- F_vs_t_curve1$ms[F_vs_t_curve1$pN == Fmax]
Fmax_index <- which(F_vs_t_curve1$pN == Fmax) # alternative: which.max(F_vs_t_curve1$pN)


```
F max = `r Fmax`, the time of F max = `r Fmax_t`.

\newpage
## Compute Contacting point

### A) Intersect between baseline and linear gradient
* We will use the force signal between time 0.0 - 0.5 ms as well as 1.5 - 2.0 ms as **baseline**.
* We will use a sliding window approach to approximate the **gradient** of the linear slope within the time 0.5 - 1.2 ms.

```{r intersect, echo=FALSE, warning=F}

## baseline
Index_t_0.5 <- nrow(F_vs_t_curve1)/4  #index after 1st quarter
Index_t_1.2 <- floor(1.2/2 * nrow(F_vs_t_curve1)) #index after 60%
Index_t_1.5 <- 3/4 * nrow(F_vs_t_curve1) #index after 3 quarters
Index_t_2.0 <- nrow(F_vs_t_curve1) #last index

Force_base_values <- c( F_vs_t_curve1$pN[1 : Index_t_0.5], F_vs_t_curve1$pN[Index_t_1.5 : Index_t_2.0])
Force_baseline <- mean(Force_base_values)

# plot baseline values and baseline (red)
Dataframe <- as.data.frame(cbind(1:(length(Force_base_values)), Force_base_values))
colnames(Dataframe) <- c("index", "force")
ggplot(data = Dataframe, mapping = aes(x = index, y = force )) +
  geom_point() +
  geom_line() +
  geom_abline(slope = 0, intercept = Force_baseline, colour = "red", lwd = 1.5) +
  ggtitle(Curve_name) +
    labs(
      x = "Time [ms]",
      y = "Force [pN]"
    )



## gradient 
# curve to consider
F_vs_t_gradient <- F_vs_t_curve1[Index_t_0.5 : Index_t_1.2, ]

# calc best fitting positive slope
library(zoo)
Zoo_data <- zoo(F_vs_t_gradient)

output <- zoo::rollapply(
  data = Zoo_data, 
  width = 8, 
  FUN = function(z) {
    lmfit <- lm(formula = pN ~ ms, data = as.data.frame(z))
    c(
      intercept = unname(lmfit$coefficients[1]),
      gradient = unname(lmfit$coefficients[2]),
      rsquared = summary(lmfit)$r.squared,
      meanms = mean(lmfit$model$ms)
    )
  },
  by.column = FALSE
)

# Filter only positive slopes
output <- as.data.frame(output)
output <- output[output$gradient > 0,]
bestfit <- as.data.frame(output)[which.max(output$rsquared),]

# plot
ggplot(data = F_vs_t_curve1, mapping = aes(x = ms, y = pN)) +
  geom_point() +
  geom_line() +
  geom_abline(intercept = bestfit$intercept, slope = bestfit$gradient, colour = "blue", lwd = 1.5) +
  ggtitle(Curve_name) +
    labs(
      x = "Time [ms]",
      y = "Force [pN]"
    )


## intersect
# bestfit$gradient * x + bestfit$intercept = Force_baseline
# x = (Force_baseline - bestfit$intercept)/ bestfit$gradient
Contacting_point_t <- (Force_baseline - bestfit$intercept)/ bestfit$gradient
Contacting_point_F <- bestfit$gradient * Contacting_point_t + bestfit$intercept 
Contacting_point   <- as.data.frame(Contacting_point_t, Contacting_point_F)

# plot
ggplot(data = F_vs_t_curve1, mapping = aes(x = ms, y = pN)) +
  geom_point() +
  geom_line() +
  geom_abline(slope = 0, intercept = Force_baseline, colour = "red", lwd = 1.5) +
  geom_abline(intercept = bestfit$intercept, slope = bestfit$gradient, colour = "blue", lwd = 1.5) +
  geom_point(data = Contacting_point, mapping = (aes(x = Contacting_point_t, y = Contacting_point_F, colour = "orange"))) +
  ggtitle(Curve_name) +
    labs(
      x = "Time [ms]",
      y = "Force [pN]"
    )

```


### B) Intersect between baseline and parabular fit & comparison to linear fit intersect
```{r parabula, echo=T}
# Audrey

```

### C) Mean of error increase from baseline (= start of adhesion dent) and error from linear gradient (= end of adhesion dent)
```{r dent, echo=T}
# suggested by Jörg Stelling

```


\newpage
## Compute **indentation depth (d)** from Contacting point for each pixel
```{r inden, echo=FALSE}

# calculate time steps between all measurements in F_vs_t_curve1
Time_difference <- Fmax_t - Contacting_point_t
Time_step <-  F_vs_t_curve1[2,1]
Time_steps_in_d <- Time_difference / Time_step

# Fmax_index  already known

# Contacting_point_index
Contacting_point_index <- unlist( Fmax_index - floor(Time_steps_in_d) )

# import distance array
Curve_Z <- "AFModulus_Flex/F_vs_Z_curves/20200619_.005.pfc-4069_ForceCurveIndex_45647.spm - NanoScope Analysis.txt"
F_vs_Z_curve <- read.table(Curve_Z, 
                            quote="\"", 
                            header = T,
                            comment.char="", 
                            stringsAsFactors=FALSE)
Distance_axis <- as.vector(F_vs_Z_curve['nm'])

# read out delta from the distance array, using the index of Fmax-time and the index of contact-point-time
d <- Distance_axis[Fmax_index, ] - Distance_axis[Contacting_point_index, ]
```
The indentation depth d = `r d` nm.

\newpage
## Compute **modulus (= stiffness, E)** for each pixel from F-max and d
```{r modulus, echo=FALSE}

# Fmax & indentation depth d were computed above

# Poisson’s ratio (typically 0.2-0.5)=0.3
v <- 0.3

# half angle of the indenter=18^0
alpha <- 180

# compute Young’s modulus
E <- (Fmax * pi * (1 - v^2)) / (2 * tan(alpha) * d^2)

```
The Young’s modulus E = `r E` [UNIT?].

\newpage
## Visualisation of the Young's Modulus
### Print picture of topology
```{r picT, echo=FALSE}

```

### Print picture of stiffness
```{r picS, echo=FALSE}

```


\newpage
## Error propagation/ sensitivity analysis of the modulus
```{r error, echo=FALSE}

```

\newpage
## Plot topology against modulus? Can this detect 'antibiotics affected areas'?
```{r further, echo=FALSE}


```

